Reproducibility of COVID-19 pre-prints in 2020

### 2021-03-05 outputs/paper/paper.Rmd
Add txt for another 1000 arxiv papers

Scientists use open repositories of papers to more quickly disseminate their research than is possible in traditional journals or conference proceedings. These repositories, such as arXiv, bioRxiv, medRxiv, and SocArXiv, are a critical component of science and many results build on the work published there. Thus it is important that the results that are published are credible. Pre-print repositories, specifically medRxiv, have drawn unprecedented attention in the context of the 2019 novel coronavirus (COVID-19) pandemic and the changes it has imposed on the scientific community [@else2020]. These repositories are not peer-reviewed, and, in general, anyone with appropriate academic credentials can submit a paper.

While neither peer-review nor credentials are a panacea nor a guarantee of quality, given the importance of these repositories, it is important that scientists impose on themselves various standards for their results. Following @weissgerber2021automated we examine papers about COVID-19 published to medRxiv through to January 2021. We search for markers of open science as indicators of reproducibility, specifically open data and open code.

@@

Our primary dataset consists of information extracted from the medRxiv repository combined with output from running a sample of COVID-19-related pre-prints through the Open Data Detection in Publications (ODDPub) text mining algorithm [@citePDDPub], using the `oddpub` package [@citeODDPubpackage].

We constructed this dataset by first creating a local copy of the medRxiv repository via the medRxiv API. We identified papers related to the COVID-19 pandemic through several keyword searches to create our sampling frame (N = 9,929, including only the most recent version of any given pre-print). This data includes the following information about each pre-print in the repository: title, abstract, author(s), date posted, research field, DOI, version number, corresponding author, corresponding author's institutional affiliation, and published DOI (if the paper has since been published in a peer-reviewed journal).

We then selected a random sample of these papers (N = 1,200) to check for open data and code markers using the ODDPub algorithm. This required downloading each paper as a PDF and then converting the PDFs to text files, using the `medrxivr` package [@citemedrxivr]. We then conducted the open data and open code detection procedure. This produced a results table indicating the presence of open data or open code markers in each paper (with a binary variable for each marker, and the relevant open data or open code statements when applicable).

Our final dataset was formed by joining these two tables together via DOI to form a dataset including all original, qualitative information for each pre-print alongside its open data and open code status and markers.

```{r monthly-papers-total, fig.cap = "Number of pre-prints related to COVID-19 posted to arXiv and medRxiv", out.width="90%", fig.align = "center"}
just_arxiv <- 
  all_arxiv_results  %>% 
  select(submitted) %>% 
  mutate(type = "arXiv") %>% 
  rename(date = submitted)

@@

both %>% 
  mutate(month = floor_date(date, "month")) %>% 
  filter(year(date) >= 2020) %>% 
  ggplot(aes(x = month)) + 
  geom_histogram(aes(fill = type), stat="count", position = "dodge") + 
  labs(title = "Pre-prints posted per month",
       subtitle = "Total of 3,379 on arXiv and 9,929 on medRxiv",
       x = "Month",
       y = "Number of pre-prints",
       fill = "Repository") +
  theme_minimal()
```

@@

@@

@@

Include info here directly from package documentation (i.e. keywords/phrases used for text parsing)?

>>>>

Scientists use open repositories of papers to more quickly disseminate their research than is possible in traditional journals or conference proceedings. These repositories, such as arXiv, bioRxiv, medRxiv, and SocArXiv, are a critical component of science and many results build on the work published there. Thus it is important that the results that are published are credible. Pre-print repositories have drawn unprecedented attention in the context of the 2019 novel coronavirus (COVID-19) pandemic and the changes it has imposed on the scientific community [@else2020]. These repositories are not peer-reviewed, and, in general, anyone with appropriate academic credentials can submit a pre-print

Neither peer-review nor credentials are a panacea nor a guarantee of quality. And the gate-keeping and slow publication times of traditional journals mean pre-print repositories have an important role to play. But given the importance of these repositories, it is important that scientists impose on themselves various standards for their results. Following @weissgerber2021automated we examine papers about COVID-19 published to  arXiv, bioRxiv, medRxiv, and SocArXiv from January 2020, through to February 2021. We search for markers of open science as indicators of reproducibility, specifically open data and open code.

@@

Our primary dataset consists of information extracted from the arXiv, bioRxiv, medRxiv, and SocArXiv pre-print repositories. We downloaded pre-prints that contained various keywords related to COVID-19 (a full list of these is available in Appendix \@ref(keywords)), and then identified the availability of open code and open data using the Open Data Detection in Publications (ODDPub) text mining algorithm [@citePDDPub] in the `oddpub` package [@citeODDPubpackage].

We constructed this dataset by first creating a local copy of the repository using the repository API. We identified papers related to the COVID-19 pandemic through several keyword searches to create our dataset of papers and associated metadata. This metadata varies by repository, but generally includes the following information: title, abstract, author(s), date posted, research field, DOI, version number, corresponding author, corresponding author's institutional affiliation, and published DOI (if the pre-prints has since been published in a peer-reviewed journal). In the case of SocArXiv there are additionally fields that are self-reported by the submitting author about the availability of code and data.

We then checked these pre-prints for open data and code markers using the ODDPub algorithm. This required downloading each pre-print as a PDF and then converting the PDFs to text files. We then conducted the open data and open code detection procedure. This involved searching for keywords and other markers of open data and open code such as references to GitHub repositories and similar. A full listing of these markers is included in Appendix \@ref(markers). This was conducted using the `open_data_search()` function from the `oddpub` package.

The result of this process is a dataset indicating the presence of open data or open code markers in each pre-print (with a binary variable for each marker, and the relevant open data or open code statements when applicable). Our final analysis dataset was formed by joining these two datasets together, typically using the DOI or the unique filename, to form a dataset including all original, qualitative metadata information for each pre-print alongside its open data and open code status and markers.

```{r monthly-papers-total, fig.cap = "Number of pre-prints related to COVID-19", out.width="90%", fig.align = "center"}
just_arxiv <- 
  all_arxiv_results  %>% 
  select(submitted) %>% 
  mutate(type = "arXiv") %>% 
  rename(date = submitted)

@@

both %>% 
  mutate(month = floor_date(date, "month")) %>% 
  filter(year(date) >= 2020) %>% 
  ggplot(aes(x = month)) + 
  geom_histogram(aes(fill = type), stat="count", position = "dodge") + 
  labs(title = "Pre-prints posted per month",
       subtitle = "Total of 3,379 relevant pre-prints on arXiv and 9,929 on medRxiv",
       x = "Month",
       y = "Number of pre-prints",
       fill = "Repository") +
  theme_minimal()
```

@@

- Add reference somewhere to that paper about the numbers: https://link.springer.com/article/10.1007/s11192-020-03675-3
- Add reference somewhere to that paper about keyword search and then analysis. Was it a Gary King one?

@@

Insert details on keyword search used to identify COVID-19 papers.

# Markers

Insert details on what markers the `oddpub::open_data_search()` uses. Include info here directly from package documentation (i.e. keywords/phrases used for text parsing)?

# Classification

@@
